---
title: "ggpipe overhead"
author: "Sergio Oller"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggpipe-overhead}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This document compares the performance of two simple plots, built with `ggplot2`
with respect to `ggpipe`, to show the performance penalty of `ggpipe`.

`ggpipe` has some performance penalty with respect to `ggplot2` because it uses
`tryCatch` and evaluates twice one of the arguments, to determine if it is a
`ggplot` object, and therefore to know if we are piping or adding.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(ggpipe)
```


We define two tasks:

- Task 1 defines a plot and saves it to a variable.
- Task 2 defines the plot and builds it.

Task 1 is where most of the overhead of `ggpipe` is supposed to be, as the plot
building step is done by `ggplot2` in any case. However Task 2 reflects the
most common use case (we usually want to print the plots we define).


Task 1 results: Define a plot

```{r task1}
b1 <- microbenchmark::microbenchmark(
  ggpipe = {p <- ggpipe::ggplot(iris) %>% ggpipe::geom_point(ggpipe::aes(x = Sepal.Length, y = Sepal.Width)) %>% ggpipe::scale_x_continuous("test")},
  ggplot2 = {p <- ggplot2::ggplot(iris) + ggplot2::geom_point(ggplot2::aes(x = Sepal.Length, y = Sepal.Width)) + ggplot2::scale_x_continuous("test")})
b1$task <- "Define Plot"

print(b1)
```

Task 2 results: Define + Build plot

```{r task2}
b2 <- microbenchmark::microbenchmark(
  ggpipe = {ggplot2::ggplot_build(ggpipe::ggplot(iris) %>% ggpipe::geom_point(ggpipe::aes(x = Sepal.Length, y = Sepal.Width)) %>% ggpipe::scale_x_continuous("test"))},
  ggplot2 = {ggplot2::ggplot_build(ggplot2::ggplot(iris) + ggplot2::geom_point(ggplot2::aes(x = Sepal.Length, y = Sepal.Width)) + ggplot2::scale_x_continuous("test"))})
b2$task <- "Define+Build Plot"

print(b2)
```


This is the median time represented in a bar plot:

```{r visualization}

b <- dplyr::bind_rows(b1, b2)

b %>%
  group_by(expr, task) %>%
  summarize(median_time = median(time)/1E6) %>%
  ungroup %>%
  ggplot %>%
  geom_col(aes(x = expr, y = median_time, fill = expr)) %>%
  xlab('Package') %>%
  ylab("Median time (ms) (less is better)") %>%
  facet_wrap(~task) %>%
  guides(fill = FALSE)

```

And the region where most of the data points are:

```{r visualization2, warning=FALSE}
ggplot(b) %>%
  geom_jitter(aes(x = expr, y = time/1E6, color = expr)) %>%
  xlab('Package') %>%
  ylab("Time (ms) (less is better)") %>%
  ylim(0, 20) %>%
  facet_wrap(~task) %>%
  guides(colour = FALSE)
```
